---
pdf_document:
  fig_caption: yes
author: "Cheng et al"
date: "October 16, 2018"
css: summarytools.css
output:
  html_document: null
  pdf_document: default
  toc: yes
geometry: margin=.25in
number_sections: yes
editor_options:
  chunk_output_type: console
title: "Luminex, Moya-moya CSF samples"
classoption: a3paper
toc: yes
---
  
# Overview: Analyze Luminex Run 1  
<div style="background-color: #d5d9e0;">
The original data input has a duplicate.
20-GS-IS18-33153172-151284-DX3	wells: D7,D6,
23-GS-IS18-33153172-151284-DX3	wells: G7,G6,
These two samples were averaged  together.  We think the tech might have run it twice because of free wells
</div>

For the final data we are going with log2 transform no normalization.  We tested two other methods outlined by Rausch et al, 2016 ( log2 follow by quantile ) and ( asinh follow by loess ) however both did not perform as well as log2 and the best performer asinh follow by loess perform very similar.  Thus to reduce data manipulation and ambiguity we are are going with just log2 not transform. 

Diffrential differences was conducted with least mean square ( LMS ) with CHEX4 as a covariated.  All p-value are corrected with tukey.   

```{r setup, include=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(include=TRUE, echo=FALSE, message=FALSE, warning=FALSE)
```


```{r, include=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
source ( "config.R")

# NOTE THERE ARE NO KEYS!  the data file contains the key 

data <- read.xlsx( data.file , sheet="remove.dup.62Plex", colNames = TRUE)
# clean up ID names 
data$ID = as.character ( sapply ( data$ID, function(x) str_match(x, ".*?\\-.*?\\-(.*?)-")[2] ) )



## create key here 
key = data[ , 1:3]

## clean up data 
df.data = t ( data[ ,5:ncol(data)] )
colnames ( df.data ) = data$ID
df.data = data.frame ( df.data )
# transform log2 
df.data.raw = df.data 
log2.p1 = log2 ( df.data + 1 )
df.data = log2 ( df.data )

## plot box plot to look at sample distribution. 

# log2 no normalize

box.raw = boxv (df.data, key, "log2 no normalize ")    

# log2 qn normalize
df.qn = data.frame ( normalize.quantiles( as.matrix ( df.data)  ) )
colnames ( df.qn ) = colnames( df.data)
row.names ( df.qn ) = row.names ( df.data )
box.qn = boxv ( df.qn, key, "log2 quantile normalize")

# asinh loess normalize 
df.asinh = asinh( df.data.raw )
df.asinh = normalize.loess( df.asinh , log.it = F)
box.df.asinh = boxv (df.asinh, key, " asinh loess normalized ")    

# OK! chose which one you wan to use! 

main = df.data
key2 = key 
file.name = "log2.noNorm"



#####
temp = melt ( main  )
colnames(temp) = c("ID", "value")
temp$group = as.character ( sapply ( temp$ID,  function(x)  unique ( key [ key$ID  == x, ]$group)  ) )

  g = ggplot(temp  , aes(value, fill=group, colours=group)) + geom_density(alpha=0.1) + 
      theme_bw()  + facet_wrap( ~ group , ncol=1, scales = "free_y"  ) 
  plot ( g )
  

# get and extract chex4 
chex = main [ grepl ( "CHEX4", row.names(main)), , drop=FALSE]
# remove all CHEX 
main = main[ !grepl ( "CHEX", row.names(main)), ]

key2$chex4 = as.numeric ( chex )

# this is corrrecting for chex4 
main2 = apply ( main, 1, function(x) x/key2$chex4)
main2 = t ( main2 )

mds.cml <-plotMDS( main  )
mds.cml <- data.frame(x=mds.cml$x, y=mds.cml$y)

m1.cml <- ggplot(mds.cml, aes(x=x, y=y, col=key2$group, shape=key2$group ) ) +
  geom_point(size=4) +
  theme_bw() +
  theme(legend.position="bottom", legend.title=element_blank(), legend.key = element_blank()) +
  labs(x="Dimension 1", y="Dimension 2") + ggtitle("") 

m1.cml 

main3 = melt ( t ( main ) )
colnames ( main3  ) = c("ID","gene","value")
main3 = merge ( main3, key2, by = "ID")

lmean = function (x ){
    bd = main3[main3$gene == x, ]
    
    bd.lm = lm(value ~ group + chex4 , data = bd)
    result = lsmeans::lsmeans(bd.lm,
        pairwise ~ group,
        adjust="tukey")

    r = summary ( result )
    return ( r$contrasts$p.value )
    
    
    
}




results = data.frame (  )
# go through each gene and analyze 
for (i in unique ( main3$gene ) ){
    l = lmean ( i )
    l = as.numeric ( l)
    
    results = rbind ( results, t( data.frame ( l )) )
}

colnames ( results )= c( "Control - MMD-hem" , "Control - MMD-ischemic", "MMD-hem - MMD-ischemic" )


results$gene = as.character ( unique ( main3$gene ) )


write.xlsx( results, paste0(out.dir, file.name , ".xlsx" ), sheetName = file.name , firstRow=TRUE, tabColour= "#ed8f55",
  col.names = TRUE, row.names = FALSE, append = FALSE)






# we have two comparisons 
key.c.mmd.hem = key2[key2$group %in% c("Control", "MMD-hem"), ] 

dge.c.mmd.hem = results[results$`Control - MMD-hem` < .05,  ]

df.temp = data.frame ( df.data[, key.c.mmd.hem$ID] )
logfc = as.numeric ( apply(df.temp, 1, function(x){
     mean ( x[33:nrow ( key.c.mmd.hem)]) -mean ( x[1:32] ) 
}) )
rtemp = results[!grepl("CHEX", results$gene), c("gene","Control - MMD-hem")]
rtemp$logFC = logfc

data.c.mmd.hem = cbind (  rtemp , df.temp  )


write.xlsx( data.c.mmd.hem, paste0("c.mmd.hem.xlsx" ), sheetName = "c.mmd.hem", firstRow=TRUE, tabColour= "#ed8f55",
  col.names = TRUE, row.names = FALSE, append = FALSE)

source('https://www.dropbox.com/s/8c2dc7aeklnniu3/general.11.R?dl=1') # get external functions


dd = df.temp [ row.names ( df.temp) %in% dge.c.mmd.hem$gene,  ]
my_palette <- colorRampPalette(c("blue", "#f2f2f2", "#fc8600"))(n = 75)
  par(mar=c(0,0,0,0))
  # bottom, left, top and right 

  hm= heatmap.2(as.matrix( dd  ),  trace="none", labRow=row.names(dd), distfun = function(x) dist(x,method = 'euclidean'), ColSideColors = this.color,
  hclustfun= function(x) hclust(x,method = 'ward.D2')  ) 
 
  this.color = ifelse ( key.c.mmd.hem$group == "Control", "blue","green")
  
   d3heatmap(as.matrix( dd  ),  ColSideColors = this.color, trace="none", labRow=row.names(dd), distfun = function(x) dist(x,method = 'euclidean'), 
  hclustfun= function(x) hclust(x,method = 'ward.D2')  )
 
## clean up name 
   key.c.mmd.hem$group = gsub ( "-| ", ".", key.c.mmd.hem$group )
   key.c.mmd.hem$tube = key.c.mmd.hem$ID
   
   gk = 'BDNF'
g = plot.gene ( df.temp [row.names(df.temp) == gk,], key.single= key.c.mmd.hem , ptext="",title= gk, pid="tube", key.c="group", add=0)
g$g2 + ylab("log2 ( x + 1) ")

data.c.mmd.hem$adj.P.Val = rep(1, nrow ( data.c.mmd.hem))
	
colnames(data.c.mmd.hem)[2] = "P.Value"


rsub = c("P.Value","adj.P.Val","logFC", "gene" )
post1 <- plot.post ( data.c.mmd.hem, "group", "group" , key.c.mmd.hem  , rsub, exp.group="group", exp.this="MMD.hem", normal.this= "Control", sample.id="tube", GENE_SYMBOL = "gene", fdr=2, fold_thres = .2, title1 = "2D vs 3D", top10 = NA)



```


                  
                  


